---
title: "Autism Studies Data"
author: "Marissa Cook"
date: '2022'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Analyzing data of autistic studies

```{r}
setwd("C://Users/maris/Downloads/Code/Autism_data")
autism_studies_stats <- read.csv("autism_prevalence_studies.csv")

```

Sort Author by alphabet

```{r}
Author_sorted <- sort(autism_studies_stats$Author)

```

Studies published in the 2000s

```{r}
studies_2000s <- subset(autism_studies_stats, Year.Published > 1999)

```

What is the average participant age in these most recent studies?

```{r}
class(studies_2000s$Age.Range)
age_range_factored<- as.factor(studies_2000s$Age.Range)
na.omit(age_range_factored)
age_range_factored <- as.data.frame(age_range_factored)
average_age_2000s <- mean(age_range_factored)

```
The structure of the data for the age range of participants used is not best suited for this analysis.


What is the average sample size?
(To remove NA values without altering initial dataset)

```{r}
sample_size <- na.omit(studies_2000s$Sample.Size)
average_n <- mean(sample_size)
print(average_n)
print(paste0(" The average sample size is ", average_n))
```

What is the average sample size of studies completed before the 2000s?

```{r}
studies_1900s <- subset(autism_studies_stats, Year.Published <2000)
sample_n_1900s <- mean(studies_1900s$Sample.Size)
print(paste0(" The average sample size for studies published before 2000 is ", sample_n_1900s))

```

Sample size comparison

```{r}
print(paste0(" The sample size for studies published in the 2000s ", (average_n), " is larger than the sample size for studies published before the 2000s ", sample_n_1900s))
```

Which autism types were analyzed in all studies?

```{r}
classified_types_ASD <- unique(autism_studies_stats$Autism.Types.Included)
print(classified_types_ASD)
```

Did the ASD prevalence estimate per 1000 individuals increase over the years?

```{r}
prevalence_2000s <- mean(studies_2000s$ASD.Prevalence.Estimate.per.1.000)
prevalence_1900s <- mean(studies_1900s$ASD.Prevalence.Estimate.per.1.000)
print(" The prevalence is reported to be higher in the 2000s.")
```

Is autism more prevalent among a specific gender?

```{r}
average_ratio <- mean(autism_studies_stats$Male.Female.Sex.Ratio)
class(autism_studies_stats$Male.Female.Sex.Ratio)
sex_ratio <- na.omit(autism_studies_stats$Male.Female.Sex.Ratio)
average_ratio <- mean(sex_ratio)
print( "Autism disorders are more prevalent among males, possibly 4x as likely.")
```

Graph number of studies over the years
Install ggplot2 from packages

```{r}
library(ggplot2)
tally_years <- table(autism_studies_stats$Year.Published)
tally_years <- as.data.frame(tally_years)
Published_Study_Progression_plot = ggplot(tally_years, aes(x=Var1, y= Freq))+
  geom_line()+
  geom_point(shape=5, color= "black", fill= "red", size=2 )+
  ggtitle("Progression of Autism Studies Published")+
  labs(x= "Years", 
       y= "Amount of Studies Published")+
  scale_x_discrete(label= function(x) stringr:: str_trunc(x,12))+
  theme(axis.text.x = element_text(angle= 90, hjust = 25, vjust = 2))

print(Published_Study_Progression_plot)
```

